---
title: "FD0000 Query Title"
description: "Please treat content as confidential"
author: Query Fulfillment Center for PCORnet®
date: last-modified
date-format: "MM-DD-YYYY"
title-block-banner: "white"
title-block-style: default
title-block-banner-color: "#272A59FF"
format:
  html:
    include-in-header:
      - text: |
           <script src="https://cdn.jsdelivr.net/npm/d3@7"></script>
           <script src="https://cdn.jsdelivr.net/npm/@observablehq/plot@0.6"></script>
           <script src="https://unpkg.com/topojson-client@3"></script>
           <script src="https://unpkg.com/topojson@3"></script>
    output-file: paqs_query_0_report
    grid: 
      body-width: 2160px
      gutter-width: 10px
      sidebar-width: 300px
      margin-width: 0px
    embed-resources: true
    smooth-scroll: true
    toc: true
    toc-location: left
    toc-depth: 3
    toc-expand: true
    fig-format: svg
backgroundcolor: "#fff"
css: ../../tools/pcornet-style-sheet.css
---


# Request Summary

#### Description


#### Planned Use


#### Recipients 

Query is sent to the participating volunteer pilot DataMarts on the most recently approved PCORnet CDM 7.0 Data Curation (or your currently approved CDM version).

#### Confidentiality

Treat request topic as confidential until final release of public report.

#### Response

Network-level aggregated data can be presented or published on public-facing websites (e.g. PCORnet.org), in a manuscript, and/or used in marketing or educational materials describing these populations in PCORnet®. DataMart (DM) response to this query is acknowledgement that data can be aggregated into a network-level report for public-facing venues.

The Coordinating Center (CC) may reuse network aggregate and DM-level data for internal planning and to inform potential funding and research opportunities.

Low cell count masking is not permitted in DataMart output returned to the CC; however, the CC will always mask low cell counts (<11) in any resulting data report to be shared with the requester.




```{r include=FALSE}
library(tibble)
library(knitr)
library(fs)
library(htmltools)
library(reactable)
library(dplyr)
library(readr)
library(purrr)
library(cohortattrition)


source('../script/base/reporting.R')
load_results <- function(name) {
    rslt <-  tryCatch(
    {read_csv(name)},
    error = function(e) {
      return(tibble())
    }
  )
  rslt
}
```

```{r include=FALSE}
rslt <- readRDS(file = "../results/fd0000_table1.rds") %>%
  prettify_table(.,cohort = TRUE, domain = TRUE,characteristics = TRUE,percentage = TRUE)

```

```{r include=FALSE}
attrition <- readRDS('../results/fd0000_attrition.rds') %>%
  as_tibble() %>%
	rename(site = any_of("qry_site")) %>% 
  prettify_table(., cohort = TRUE)
```

```{r include=FALSE}
attrition <- reduce(lapply(split(attrition, attrition$Cohort), function(x) x %>% cohortattrition::ca_process(multi_or_single_site = "single",anomaly_or_exploratory = "exploratory",start_step_num = 1) ), bind_rows)
```

```{r include=FALSE}
signature <- load_results('../results/signature.csv') %>%
  mutate(
    `Total Run Time (in minutes)` = `Query End Time` - `Query Start Time`
  ) %>%
  select(-`Total Run Time`)

```

## Query Execution Time
```{r}
#| message: false
#| warning: false
#| errors: false
#| echo: false

kable(signature)

```

&nbsp;


```{r}
#| message: false
#| warning: false
#| errors: false
#| echo: false

browsable(observable_plots_table(rslt,attrition))

```